---
title: "Bayesian Modeling and Prediction for Movies Dataset"
output: 
  prettydoc::html_pretty: 
    fig_height: 6
    fig_width: 9
    highlight: github
    theme: tactile
    toc: yes
    df_print: kable
    keep_md: true
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
library(BAS)
```

### Load data



```{r load-data}
load("movies.Rdata")
```



* * *

## Part 1: Data

In this study random sample were used and results can be generalized to the US population. Since no random assignment was used so causality can not be inferred. 

* * *

### Research Queastion

Can we predict audience score from variable that are mentioned in project direction?

## Part 2: Data manipulation

lets explore the data 

```{r}
 
summary(movies)
```

Now create new variable: 

```{r}

movies <- movies %>% 
    mutate(feature_film = ifelse(title_type == 'Feature Film', 'yes', 'no'),
           drama =        ifelse(genre =='Drama', 'yes', 'no'),
           mpaa_rating_R =  ifelse(mpaa_rating == 'R', 'yes', 'no'),
           oscar_season = ifelse(thtr_rel_month %in% c(10:12), 'yes', 'no'),
           summer_season = ifelse(thtr_rel_month %in% c(5:8), 'yes','no'))

```

explore the dataset again.

```{r}
glimpse(movies)
```

Now convert the newly created variable to factor type from character type.

```{r}


movies[c('feature_film',
         'drama', 'mpaa_rating_R', 
         'oscar_season',
         'summer_season')] <- lapply(movies[c('feature_film', 
                                                'drama', 'mpaa_rating_R', 
                                                'oscar_season', 
                                              'summer_season')], factor)
```


```{r}
glimpse(movies)
```



Remove the NA value from runtime since we are going to use this as predictor later on.

```{r}
movies <- filter(movies, !is.na(runtime))
```


* * *

## Part 3: Exploratory Data Analysis

### Drama

```{r}

    ggplot(data = movies, mapping = aes(x = drama ,
                                        y = audience_score, 
                                        fill = drama))+
    geom_boxplot(outlier.colour = 'red',
                 outlier.shape = 1,
                 notch = F,
                 width = 0.5)+
    xlab('Drama')+
    ylab('Audience Score')+
    ggtitle('Difference in Audience Score in Drama' )+
    theme(panel.background = element_rect(fill = 'light blue'))+
    theme(plot.title = element_text(hjust = 0.5))

```

Mean audience score is higher for drama movies.

```{r}

select(movies, audience_score, drama) %>% 
    group_by(drama) %>% 
    summarise(audience_score_mean = mean(audience_score),
              min = min(audience_score),
              max = max(audience_score))
```

Mean audience score is 65.3 for drama category which is higher than movie that is not drama.


### Feature Films

```{r}
select(movies, audience_score, feature_film) %>% 
    group_by(feature_film) %>% 
    summarise(audience_score_mean = mean(audience_score),
              min = min(audience_score),
              max = max(audience_score))
```

Feature Film has lower mean audience score which is 60.5.

```{r}

    ggplot(data = movies, mapping = aes(x = feature_film ,
                                        y = audience_score, 
                                        fill = feature_film))+
    geom_boxplot(outlier.colour = 'red',
                 outlier.shape = 1,
                 notch = F,
                 width = 0.5)+
    xlab('Feature Film')+
    ylab('Audience Score')+
    ggtitle('Difference in Audience Score in Feature Film' )+
    theme(panel.background = element_rect(fill = 'light blue'))+
    theme(plot.title = element_text(hjust = 0.5))
```

Film that is not featured has higher mean audience score.

### mpaa_rating_R

```{r}
select(movies, audience_score, mpaa_rating_R) %>% 
    group_by(mpaa_rating_R )%>% 
    summarise(audience_score_mean = mean(audience_score),
              min = min(audience_score),
              max = max(audience_score))
```

very little difference in mean audince score whether mpaa rating is R or not.


```{r}
 ggplot(data = movies, mapping = aes(x = mpaa_rating_R , 
                                     y = audience_score,
                                     fill = mpaa_rating_R))+
    geom_boxplot(outlier.colour = 'red',
                 outlier.shape = 1,
                 notch = F,
                 width = 0.5)+
    xlab('mpaa_rating_R')+
    ylab('Audience Score')+
    ggtitle('Difference in Audience Score in mpaa_rating_R' )+
    theme(panel.background = element_rect(fill = 'light blue'))+
    theme(plot.title = element_text(hjust = 0.5))
```

We can also visualize the small differnce.

### Oscar Season

```{r}
select(movies, audience_score, oscar_season) %>% 
    group_by(oscar_season )%>% 
    summarise(audience_score_mean = mean(audience_score),
              min = min(audience_score),
              max = max(audience_score))
```


mean audince score is 63.6 when movie release in oscar season which is higher.

```{r}
 ggplot(data = movies, mapping = aes(x = oscar_season , 
                                     y = audience_score, 
                                     fill = oscar_season))+
    geom_boxplot(outlier.colour = 'red',
                 outlier.shape = 1,
                 notch = F,
                 width = 0.5)+
    xlab('Oscar Season')+
    ylab('Audience Score')+
    ggtitle('Difference in Audience Score in Oscar Season' )+
    theme(panel.background = element_rect(fill = 'light blue'))+
    theme(plot.title = element_text(hjust = 0.5))
```

difference can be visualize and we can see that difference is small.


### Summaer Season

```{r}
select(movies, audience_score, summer_season) %>% 
    group_by(summer_season )%>% 
    summarise(audience_score_mean = mean(audience_score),
              min = min(audience_score),
              max = max(audience_score))
```

Mean audience score for summer season 61.8 which is lower than not release in summer season.


```{r}
 ggplot(data = movies, mapping = aes(x = summer_season,
                                     y = audience_score, 
                                     fill=summer_season))+
    geom_boxplot(outlier.colour = 'red',
                 outlier.shape = 1,
                 notch = F,
                 width = 0.5)+
    xlab('Summer Season')+
    ylab('Audience Score')+
    ggtitle('Difference in Audience Score in Summer Season' )+
    theme(panel.background = element_rect(fill = 'light blue'))+
    theme(plot.title = element_text(hjust = 0.5))
```

from the above plot we can see that movie that is released in summer season has lower mean audience score.


* * *

## Part 4: Modeling

create a dataframe with required variaables.

```{r}
movies <- select( movies,
                  audience_score,
                 feature_film, 
                 drama,
                 runtime,
                 mpaa_rating_R,
                 thtr_rel_year,
                 oscar_season,
                 summer_season,
                 imdb_rating,
                 imdb_num_votes,
                 critics_score,
                 best_pic_nom,
                 best_pic_win,
                 best_actor_win,
                 best_actress_win,
                 best_dir_win,
                 top200_box)
```


explore the dataset.

```{r}
summary(movies)
```

### Model


```{r}
model_bayes2 <- bas.lm(audience_score ~ ., 
                       data=movies,
                       method='MCMC', 
                       prior='ZS-null',
                       modelprior=uniform())
```



### Diagnostics


```{r}
plot(model_bayes2)
```

```{r}
image(model_bayes2, rotate = F)
```


```{r}
diagnostics(model_bayes2, type = 'model')
```

```{r}
round(summary(model_bayes2),3)
```


### Coefficients

We can plot the the important predictor ibdb raing and critics score from above image.

```{r}
par(mfrow = c(1,3))
plot(coefficients(model_bayes2), subset = c(1,9,11), ask = FALSE)
```



* * *

## Part 5: Prediction


```{r}



beast_find <- data.frame(audience_score=79,
                         feature_film=factor("yes", levels=c("yes", "no")),
                         drama=factor("no", levels=c("yes", "no")),
                         runtime=132,
                         mpaa_rating_R=factor("no", levels=c("yes", "no")),
                         thtr_rel_year=2016,
                         oscar_season=factor("yes", levels=c("yes", "no")),
                         summer_season=factor("no", levels=c("yes", "no")),
                         imdb_rating=7.4,
                         imdb_num_votes=320961,
                         critics_score=74,
                         best_pic_nom=factor("no", levels=c("yes", "no")),
                         best_pic_win=factor("no", levels=c("yes", "no")),
                         best_actor_win=factor("no", levels=c("yes", "no")),
                         best_actress_win=factor("no", levels=c("yes", "no")),
                         best_dir_win=factor("no", levels=c("yes", "no")),
                         top200_box=factor("no", levels=c("yes", "no")))
                         
```




```{r}
score_predic<- predict(model_bayes2, 
                       newdata=beast_find, 
                       estimator="BMA", 
                       se.fit=TRUE)

```

```{r}
#margin

margin <- qt(0.95, df=score_predic$se.bma.pred[1]) *
                     mean(score_predic$se.bma.pred)

```

```{r}
beast_predic_interval <- data.frame(
    'title'      ="Fantastic Beast and where to find them",
    'prediction' =sprintf("%2.1f", score_predic$Ybma),
    'interval'   =sprintf("%2.1f - %2.1f", score_predic$Ybma - margin,
                          score_predic$Ybma + margin),
    
    'actual'     =79)

beast_predic_interval
```



### Reference(s) 

**Movie** - **FANTASTIC BEASTS AND WHERE TO FIND THEM (2016)**   
**data source link** -
[imdb](https://www.imdb.com/title/tt3183660/?ref_=nv_sr_2) and   [rottentomatoes](https://www.rottentomatoes.com/m/fantastic_beasts_and_where_to_find_them)

* * *

## Part 6: Conclusion

Important predictor are imdb rating and critics score. 