---
title: "Statistical Inference with the GSS Data"
output: 
  prettydoc::html_pretty:
    fig_height: 4
    highlight: pygments
    theme: cayman
    toc: yes
    df_print: kable
    keep_md: true
---


### Load packages

```{r load-packages, message = FALSE}
library(tidyverse)
library(statsr)
```

### Load data



```{r load-data}
load("gss.Rdata")
```



* * *

## Part 1: Data

Random saple were used in this survey. Resutls can be generalized to US population. No random assignment was used so causality cannot be inferred.

* * *

## Part 2: Research Question

There is a general idea that women are more compassionate than men.  So there is a chance that attitude to some social problem would be different between men and women.  
Now, I want to see is there any difference in attitude of men and women toward the spending of the money for improving conditions of the blacks. 

variable to be explored-  

* sex - respondent's sex  
* natrace - attitude toward spending money for improving the conditions of blacks.  

I will use data from 2012 for my ressearch question.

* * *

## Part 3: Exploratory data analysis

I will build the contingency table for my two categorical variable sex and natrace.  
Before that I will build a dataframe with the variable of interest. 

```{r subset_columns}
# using select command from dplyr to build a dataframe with selected columns.
gss <- select(gss,sex, natrace, year)

```

Since I am only interested in 2012 data so I will filter the data using filter command from dplyr and I will also drop all missing values(NA).

```{r filter_2012 data}
gss <- gss %>% 
    filter(year ==  2012) %>% 
    drop_na()
    
```

Lets check the structure of the data


```{r data structure}
# structure of the data
str(gss)
```


```{r dimension of data}

# dimension of data

dim(gss)

```

```{r head of data}

# head of data

head(gss)

```

```{r summary}

# quick summary
summary(gss)
```


Now build the contingency table to summarise my data.  

```{r make_table}

# I can use table function like this 
gss_tab <- table(gss$sex, gss$natrace)
gss_tab
```

OR

```{r xtable}
# I can use xtabs function like this
gss_xtab <- xtabs(~sex + natrace, gss)
gss_xtab

# I will use gss_xtab table for further analysis
```

I can add column sum and row sum to my table if I want,    
using addmargins command.  

```{r addmargins}
addmargins(gss_xtab)
```

Now lets convert to proportion table.   

```{r proportion_table}

#proportion to row sum and will round the value to 3 digits
round(prop.table(gss_xtab, margin = 1), 3)
```

I want to see as percentage so I have to multiply by 100 here.  

```{r percentage}

# as percentage
100 * (round(prop.table(gss_xtab, margin = 1), 3))
```

From the above table, 32% male  think that too little money is spending for improving
conditions of blacks where 42% female think as same.   Also there are diffence in male and female percentage who think spending is about right or too much.

Now plot the data to visualize.  

```{r plot, fig.align='center'}
ggplot(data = gss, aes(x= natrace, fill=sex))+
    geom_bar()+
    xlab('Money Spending')+
    ylab('Count')+
    labs(fill = 'Sex')+
    ggtitle('Attitudinal Differences of Men and Women')+
    theme(plot.title = element_text(hjust = 0.5))
```

From the above bar plot,    
there are difference in male and female proportion in each sub categoty of spending money.



* * *

## Part 4: Inference

### Hypothesis

Null Hypotheis:  

Sex and attitude toward spending for improving conditions of blacks are independent.

Alternative Hypothesis:  

Sex and attitude toward spending for improving conditions of blacks are dependent.

### Conditions

I am using two categorical varible with two or more levels, so we will use chi-square test as statistical method. Now check whether our data met other conditions for chi-square test.

* Random sample -- our sample are random
* if sampling without replacement, n< 10 % of population -- sample is less than 10% of population, n is 877.
* each case only contributes to one cell ine the table -- each case only contributed once.
* sample size: each particular scenario must have at least 5 expected case -- each cell have more than 5 cases.



### Statistical Method

As I mentioned before since I am using two categorical variable with two or more levels I am going to use chi-square test as my statistical method to see wheter there is association between sex and attitude toward spending money for blacks.



### Code and output

```{r chi-square test}
# I have used gss_xtab table that I have built earlier

chisq.test(gss_xtab)

```



### Interpretations

From the output, the p value is 0.02201.  
Since p value is less than 0.05, we reject the null hypothesis.   
So, Conclusion is that Sex and attitude toward spending for improving conditions of blacks are dependent.


### Reason for why Cl in not included


Confidence Interval is not included because both Variable are categorical with two or more variables and chi-square test has been used.
