---
title: "Probability Distribution"
output: github_document

editor_options: 

  chunk_output_type: inline

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Binomial Distribution
Hospital records show that of patients suffering from a certain disease, 75% die of it. What is the probability that of 6 randomly
selected patients,4  will recover?


\usepackage{amsmath}
$PMF  :   {{n}\choose{k}} \cdot p^k(1-p)^{n-k}$


here, die = 75%
so p= .25
n= 6
k=4
 
```{r}
dbinom(4, size = 6, prob = .25)
```
 
 
 plot the distribution
 
```{r}
library(tidyverse)
y <- dbinom(0:6, 6, .25)
y
x <- c(0:6)
x
plot(x,y, type = 'h', col='blue')
```
 
 
 

### Bernoulli Distribution

### Geometric Distribution

### Hypergeometric Distribution

### Negative binomial Distribution

### Poisson Distribution

### Normal Distribution

### Exponential Distribution

### Beta Distribution

### Gamma Distribution

### Cauchy Distribution

### chi-squared 

### Erlang Distribution

### F distribution

### Student's t distribution




question: 




